---
title: "exercise-08"
author: "Yen Do"
format: html
editor: visual
---

## Step 1: Loading data

```{r}
library(tidyverse)
library(skimr)
f <- ("https://raw.githubusercontent.com/difiore/ada-datasets/main/Street_et_al_2017.csv")
d <- read.csv(f, header = TRUE)
skim(d)
```

## Step 2: Plot

```{r}
ecv <- d$ECV #ECV
gs <- d$Group_size #social group size
long <- d$Longevity #Longevity
jpl <- d$Weaning #juvenile period length
rl <- d$Repro_lifespan #reproduction lifespan
par(mfrow = c(2,2))
plot (data = d, ecv ~ long)
plot(data = d, ecv ~ gs)
plot(data = d, ecv ~ jpl)
plot(data = d, ecv ~ rl)
```

## Step 3:

```{r}
d1 <- select(d, ECV = "ECV", Social_group_size = "Group_size")
d1 <- na.omit(d1)
beta1 <- cor(d1$ECV, d1$Social_group_size) * (sd(d1$ECV)/sd(d1$Social_group_size))
beta1
beta0 <- mean(d1$ECV) - beta1 * mean(d1$Social_group_size)
beta0
```

## Step 4:

```{r}
m <- lm(data = d1, ECV ~ Social_group_size)
m
```

## Step 5:

```{r}
catarrhinines <- filter(d, Taxonomic_group == "Catarrhini")
m_cata <- lm(data = catarrhinines, ECV ~ Group_size)

platyrrhinines <- filter(d, Taxonomic_group == "Platyrrhini")
m_platy <- lm(data = platyrrhinines, ECV ~ Group_size)

strepsirhines <- filter(d, Taxonomic_group == "Strepsirhini")
m_strep <- lm(data = strepsirhines, ECV ~ Group_size)

m_cata
m_platy
m_strep
```

Answer: Yes, the regression of coefficients differ among groups based on slop beta01. If β1 values are very different among groups, their relationships will be different.

## Step 6: Caculate SE for slope coefficient, 95% CI and p value and confirm by lm() function

```{r}
residuals <- d1$ECV - (beta0 + beta1 * d1$Social_group_size)
s <- sqrt(sum(residuals^2) / (nrow(d1) - 2))

#  SE for beta1
SEbeta1 <- s / sqrt(sum((d1$Social_group_size - mean(d1$Social_group_size))^2))
SEbeta1


# 95% CI
t_value <- qt(0.975, df = nrow(d1)-2)
CI_lower <- beta1 - t_value * SEbeta1
CI_upper <- beta1 + t_value * SEbeta1

# p-value
t_stat <- beta1/SEbeta1
p_value <- 2*pt(-abs(t_stat), df = nrow(d1) -2)

SEbeta1
CI_lower
CI_upper
p_value

#confirm with lm() function
m <- lm(formula = ECV ~ Social_group_size, data = d1)
SEbeta1_lm <- summary(m)$coefficients["Social_group_size", "Std. Error"]
SEbeta1_lm
CI_lm <- confint(m, level = 0.95)["Social_group_size", ]
CI_lm
p_value_lm <- summary(m)$coefficients["Social_group_size", "Pr(>|t|)"]
p_value_lm
broom::tidy(m)
```

# Step 7: Permutation test

```{r}
library(infer)

# Set the number of permutations
nperm <- 1000

# Generate permutation distribution for slope
perm <- d1 |>
  specify(ECV ~ Social_group_size) |> 
  hypothesize(null = "independence") |> 
  generate(reps = nperm, type = "permute") |> 
  calculate(stat = "slope")  

# Calculate the observed slope from the original data
observed_slope <- lm(ECV ~ Social_group_size, data = d1) |>
  broom::tidy() |>
  filter(term == "Group_size") |>
  pull(estimate)

observed_slope

# Calculate the p-value for the observed slope  using the quantile method
p_value_perm <- mean(abs(perm$stat) >= abs(observed_slope))

p_value_perm

```

Answer: In this permutation test, I permute the ECV values while keeping Social_group_size (or Group_size) fixed. The p-value can tell us how extreme our observed slope is compared to a permutation. Normally, if p \< 0.05, we can tell that two variables are significantly associated. From our permutation test, p-value = 0, it means the social group size is significantly associated with brain size (ECV).

## Step 8: Bootstrap Confidence Intervals

```{r}
boot_slopes <- replicate(1000, {
  d_boot <- d1 |> sample_frac(replace = TRUE)
  coef(lm(ECV ~ Social_group_size, data = d_boot))[2]
})
boot_CI_quantile <- quantile(boot_slopes, probs = c(0.025, 0.975))
boot_SE <- sd(boot_slopes)
boot_CI_theory <- c(beta1 - 1.96 * boot_SE, beta1 + 1.96 * boot_SE)

# Print results
boot_CI_quantile
boot_CI_theory
```

Answer: Yes. These CIs from both quantile and theory-based method don't contain 0. It suggests that the slope (beta1) is significant different from 0.
